Tropical Andes Functional diversity for plants and Frugivores

### 2024-06-03 code/L2/L2_1_TropicalAndes_funcdiv.Rmd
trying to fix functional diversity mapping by gridding point fdis data - produced map but doesn't look like 100km

```{r}
# Create a grid over the area
create_grid <- function(data, cellsize) {
  st_make_grid(st_as_sfc(st_bbox(data)), cellsize = cellsize) %>%
    st_sf() %>%
    mutate(cellid = row_number())
}

# Aggregate functional dispersion within each grid cell
aggregate_to_grid <- function(grid, points, value_col) {
  points %>%
    st_join(grid) %>%
    group_by(cellid) %>%
    summarize(mean_value = mean(get(value_col), na.rm = TRUE)) %>%
    ungroup() %>%
    st_set_geometry("geometry")
}

# use funtion
resolution_meters <- 100000
variable_name <- "fdis_frugivore"

# Create grid
grid <- create_grid(spatial_fdis_frugivore, cellsize = resolution_meters)

# Aggregate data to grid
aggregated_data <- aggregate_to_grid(grid, spatial_fdis_frugivore, variable_name)

# Convert geometries to centroids
aggregated_data_centroids <- st_centroid(aggregated_data)

# Extract coordinates from centroids
aggregated_data_coords <- st_coordinates(aggregated_data_centroids)

# Add coordinates as separate columns
aggregated_data <- aggregated_data %>%
  mutate(x = aggregated_data_coords[, 1], y = aggregated_data_coords[, 2])

# Remove rows with NA values in mean_value
aggregated_data <- aggregated_data %>%
  filter(!is.na(mean_value))
```

```{r}
FDisplot <-
    ggplot() +
    geom_sf(data = Americas, fill = "white")+
    geom_sf(data = TApoly, fill = "lightgrey", size = 0.1) +
    geom_sf(data = grid, fill = "grey")+
    geom_sf(data = aggregated_data, aes(fill = mean_value), color = NA) +
    labs(x = "Longitude", y = "Latitude", fill = "FDis") +
    scale_fill_gradient(low = "blue", high = "red") +
    coord_sf(xlim = c(-82, -60), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
    theme(panel.background = element_rect(fill = "lightblue")) +
    theme(text = element_text(size = 12)) # Explicitly setting text size
FDisplot
```

# Mapping without projection 
Because projecting is being difficult
```{r}
TropicalAndes_IUCNHabitat_Forest_latlong <- read_sf(file.path(data_path_L0, "Forest_sf.shp"), layer = "Forest_sf")
Americas_latlong <- ne_countries(continent = c("North America", "South America"), returnclass = "sf")
#polygon of Tropical Andes
TApoly_latlong <- Americas_latlong %>% filter(sovereignt == "Bolivia" |sovereignt == "Ecuador" | sovereignt == "Venezuela" | sovereignt == "Colombia" | sovereignt == "Peru")
```

```{r}
# Generate coordinates
subset_coords_frugivore <- site_loc_key_frugivore[rowSums(PAM_frugivore_site_final) >= 5, ]
subset_coords_frugivore_sp <-subset_coords_frugivore[,1:2 ]

```{r}
# Convert the dataframe to sf format
spatial_fdis_frugivore <- st_as_sf(frugivore_fd_sp, coords = c("Longitude.x.", "Latitude.y."))
```

```{r}
# set crs of sf objects
spatial_fdis_frugivore <- spatial_fdis_frugivore %>% st_set_crs(4326)
```

```{r}
# Create a grid over the area
create_grid <- function(data, cellsize) {
  st_make_grid(st_as_sfc(st_bbox(data)), cellsize = cellsize) %>%
    st_sf() %>%
    mutate(cellid = row_number())
}

# Aggregate functional dispersion within each grid cell
aggregate_to_grid <- function(grid, points, value_col) {
  points %>%
    st_join(grid) %>%
    group_by(cellid) %>%
    summarize(mean_value = mean(get(value_col), na.rm = TRUE)) %>%
    st_set_geometry("geometry")
}

# Example usage with your data
resolution_meters <- 100000
variable_name <- "fdis_frugivore"

# Create grid
grid <- create_grid(spatial_fdis_frugivore, cellsize = resolution_meters)

# Aggregate data to grid
aggregated_data <- aggregate_to_grid(grid, spatial_fdis_frugivore, variable_name)

```{r}
FDisplot <-
    ggplot() +
    geom_sf(data = Americas_latlong, fill = "white")+
    geom_sf(data = TApoly_latlong, fill = "lightgrey", size = 0.1) +
    geom_point(data = spatial_fdis_frugivore , aes(fill = fdis_frugivore)) +
    scale_fill_gradient(low = "blue", high = "red") +
    coord_sf(xlim = c(-82, -60), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
    theme(panel.background = element_rect(fill = "lightblue")) +
    annotation_scale(location = "bl", width_hint = 0.5) +
    labs(x = "Longitude", y = "Latitude", fill = "FDis")
FDisplot
```

# Plants
Test with small random sample as a subset to make it work before using all data
```{r}
set.seed(100)
all_plant_species <- row.names(plant_traits_df_final)
plant_random_subset <- sample(all_plant_species, 1000)
# objects that need to be subset
PAM_plant_site_final_df <- as.dataframe(PAM_plant_site_final) #matrix
plant_traits_df_final
```

## Create trait type table
```{r}
trait_name <- c("Growth.Form", "Plant.Height_m", "Fruit.Length_mm", "Fruit.Type")
trait_type <- c("N", "Q", "Q", "N") 
plant_trait_cat <- as.data.frame(cbind(trait_name, trait_type))
```

@@

## Create an empty dataframe to store results
plant_raos_q_results <- data.frame(spatial_scale = numeric(),
                             raos_q = numeric())

>>>>

```{r}
# Extract coordinates 
spatial_fdis_frugivore_coords <- st_coordinates(spatial_fdis_frugivore)

# Add coordinates as separate columns
spatial_fdis_frugivore <- spatial_fdis_frugivore %>%
  mutate(x = spatial_fdis_frugivore_coords[, 1], y = spatial_fdis_frugivore_coords[, 2])
```

```{r}
summary(spatial_fdis_frugivore)
```

```{r}
TAGrid_100km <- TApoly %>%
  st_make_grid(cellsize = c(100000)) %>%
  st_intersection(TropicalAndes_IUCNHabitat_Forest) %>%
  st_cast("MULTIPOLYGON") %>%
  st_sf() %>%
  mutate(cellid = row_number())

spatial_fdis_frugivore$cellid <- sapply(1:nrow(TAGrid_100km), function(i) {
  contains <- st_contains(TAGrid_100km[i, ], spatial_fdis_frugivore)
  cellid <- ifelse(any(contains), i, NA)
  return(cellid)
})

spatial_fdis_frugivore_grid <- TAGrid_100km %>%
   st_join(spatial_fdis_frugivore) %>%
   group_by(cellid) %>%
   summarize(fdis_value = sum(fdis_frugivore, na.rm = TRUE)) %>%
   ungroup()
```

```{r}
summary(spatial_fdis_frugivore_grid)
```

```{r}
FDisplot <-
    ggplot() +
    geom_sf(data = Americas, fill = "white")+
    geom_sf(data = TApoly, fill = "lightgrey", size = 0.1) +
    geom_sf(data = spatial_fdis_frugivore_grid, aes(fill = fdis_value), color = NA) +
    labs(x = "Longitude", y = "Latitude", fill = "FDis") +
    scale_fill_distiller(palette = "YlOrBr", direction = 1) +
    coord_sf(xlim = c(-82, -60), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
    theme(panel.background = element_rect(fill = "lightblue"),
        text = element_text(size = 12))
FDisplot
```

# Plants
## Create trait type table
```{r}
trait_name <- c("Growth.Form", "Plant.Height_m", "Fruit.Length_mm", "Fruit.Type")
trait_type <- c("N", "Q", "Q", "N") 
plant_trait_cat <- as.data.frame(cbind(trait_name, trait_type))
```

@@

```{r}
## Create an empty dataframe to store results
plant_raos_q_results <- data.frame(spatial_scale = numeric(),
                             raos_q = numeric())
